---
title: "Motor Trend Report"
author: "P.Y.Fong"
date: "12/31/2020"
output:
  pdf_document:
    fig_width: 10
fontsize: 10pt
geometry: "left=2cm,right=2cm,top=.5cm,bottom=.5cm"
---
```{r setup, include=FALSE}
library (knitr)
opts_chunk$set(cache =T,
               echo = T,
               message=F,
               warning=F)
```
## 1.0    Executive Summary  
In this report, the data set `mtcars` extracted from the 1974 *Motor Trend* US magazine, was used to explore the relationship between a set of predictor variables and the response variable, fuel consumption (MPG).  
The data set comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models). [*Link*](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html)  
Statistical Inference & Linear Regression methods were used to answer the following two questions:  
1.    “Is an automatic or manual transmission better for MPG?”  
2.    "Quantify the MPG difference between automatic and manual transmissions"  

The results of the analysis show that Manuals are significantly better for MPG.  
However, MPG is negatively correlated to vehicle weight and positively correlated to 1/4 mile times regardless of transmission type.
The model that explains the relationship between these variables is:  
    **MPG = 9.6 - 3.9(wt) + 1.2(qsec) + 3 (am: 0 for autos & 1 for manual)**  
On the average Manuals are approx. 7 mpg better than Automatics in the data set.  
Residuals analysis shows further exploration is needed and should consider a non linear model, normalised data values and certain outlier data values could be excluded.  

## 2.0    Loading Libraries & Data.  
```{r Load}
library(ggplot2); library(GGally); data(mtcars); attach(mtcars)
```
## 3.0    Exploratory Data Analysis.
```{r Explore}
str(mtcars) # Data set structure
```
Some variables were converted from numeric to factors.
```{r Factor}
mtcars$cyl <- as.factor(mtcars$cyl); mtcars$vs <- as.factor(mtcars$vs);
mtcars$am <- as.factor(mtcars$am); mtcars$gear <- as.factor(mtcars$gear);
mtcars$carb <- as.factor(mtcars$carb)
```
### 3.1 Exploratory Observations.  
1.    (Refer A-1) Paired graph show high correlations between "mpg" with "disp", "hp", "wt" and "qsec". The predictors mentioned were also highly correlated among themselves.
2.    The paired relationship between "mpg" and "am" warranted a closer look. (Refer A-2) Box plot, Manuals have higher values of MPG in general compared to Automatics.

##  4.0   Analysis with two sample T-test.  
Null hypothesis: Mean MPG(auto) = mean MPG(manual), assuming normal distribution.  
```{r t-test}
t.test(mpg ~ am)
```  
The null hypothesis was rejected. P-value < 0.05 (alpha = 5%). i.e. Mean MPG(auto) and mean MPG(manual) are significantly different. Also note that mean consumption (MPG) was about 7.2 mpg better for manuals.

## 5.0    Regression Analysis.  
Using backward regression model fitting, a series of models were fitted with results (refer A-3) as follows:-  

1.    **fit 10. The all variable model.**  
      Weak coefficient significance for all predictor variables i.e. p value > 0.05.  
      F stat score 7.83. The model is not statistically significant.  
      Adjusted R-squared = 0.0.779. The model explained approx 78% of the mpg variable. 

2.    **fit6. The ignore factors model** factor variables cyl, vs, gear and carb removed.
      Coefficient significance improved for wt and qsec. P value > 0.05 remain for others.  
      F stat score 27.09. The model is still not statistically significant.  
      Adjusted R-squared = 0.8347. The model explained approx. 83% of the mpg variable. 

3.    **fit5. Reduced model** variable drat removed based on coefficient abs.t values < 1.0  
      Coefficients of variables wt, qsec and am become significant. P value > 0.05 for others.  
      F stat score 32.96. The model is still not statistically significant.  
      Adjusted R-squared 0.8375. Slight improvement, explaining approx. 84% of the mpg variable. 

4.    **fit3. Final model** variables disp and hp removed based on coefficient t values < 2.0  
      Coefficients significant for predictor variables wt, qsec and am. P values < 0.05.  
      F stat score 52.75. The model is somewhat statistically significant.  
      Adjusted R-squared 0.8336 ie the model explains approx 84% of the mpg variable.   

## 6.0    Model Scatter Plots. (refer A-4 & A-5)  
The scatter plots for the final model fit3 show that MPG is negatively correlated to weight (wt) and positively correlated to 1/4 mile times (qsec) with manuals better at mpg in general.  

## 7.0    Residuals Analysis. (refer A-6)  
The Residuals vs Fitted plot shows a pattern suggesting that the linear model is not ideal.  
The Q-Q plot confirms the assumption of a normally distributed data set.  
The Scale-Location plot shows a large spread and an angled fit line, so the assumption of equal variance is weak.  
The Residuals vs Leverage shows that the model is influenced by a few outliers. 

\pagebreak

##    Appendices
###   A-1. Paired Graph
```{r paired, fig.height= 7}
ggpairs(mtcars, aes(colour=am),axisLabels ="none", lower=list(combo=wrap("facethist", binwidth=10)))
``` 

###   A-2. Boxplot of MPG vs. Transmission.
```{r boxplot,fig.height= 4}
ggplot(mtcars, aes(x=am, y=mpg, fill=am))+ geom_boxplot()+ 
  theme(axis.title.x= element_blank(),legend.position="none")+
  scale_x_discrete(labels = c('Manual','Automatic'))
```  
\pagebreak

###   A-3. Regression Model Fitting  
```{r, regression}
fit10 <-  lm(mpg ~ ., data=mtcars); summary(fit10)            # all variables
fit6 <-  update(fit10,. ~ .-cyl-vs-gear-carb); summary(fit6)  # less cyl,vs, gear & carb
fit5 <-  update(fit6,. ~ .-drat); summary(fit5)               # less gear& drat
fit3 <-  update(fit5,. ~ .-disp-hp); summary(fit3)            # less disp & hp
```  
\pagebreak

###   A-4.    Scatter Plot of MPG vs. Weight by Transmission  
```{r mpg vs wt, fig.height= 5}
ggplot(mtcars, aes(x=wt, y=mpg, group=am, color=am))+ geom_point()+
  stat_smooth(method = "lm")+ xlab("Weight")+
  scale_colour_discrete(labels=c("Automatic", "Manual"))
```

###   A-5.    Scatter Plot of MPG vs. Quarter Mile Time  
```{r mpg vs qsec, fig.height= 5}
ggplot(mtcars, aes(x=qsec, y=mpg, group=am, color=am))+ geom_point()+
  stat_smooth(method = "lm")+ xlab("Weight")+
  scale_colour_discrete(labels=c("Automatic", "Manual"))
```
\pagebreak

###   A-6.    Residual Plots
```{r residuals, fig.height=10}
library (ggfortify); autoplot(fit3)
```  